Crowd density estimation from Wi-Fi positioning data in the Amsterdam Arena

### 2016-06-28 PaperOnSmoothing/draft-31052016.tex
Updated methods section.

The data provides us with a series of estimated positions, together with corresponding values for the uncertainty (see Section on Positioning).
From the estimated positions we wish to generate density histograms.

To make density estimates we consider a two-dimensional binned region of space, and count the number of positions that fall into each bin.
We wish to visualize / preserve the uncertainty for each estimated position.
Therefore we transform each bin count into a bivariate Gaussian, using the separate error values ($\sigma_{x}$ and $\sigma_{y}$) for each dimension / direction.
For each MAC we construct a two-dimensional density function that is the sum of these smooth positions.
Because the number of measurements are different for each detected MAC address (in the interval) we normalize each histogram separately.

We sum all the density histograms to create the total probability density histogram, which should ideally integrate to the number of MAC's detected (in the interval).

We smooth each position by evaluating the following function at each bin in the histogram,

\begin{equation}
\hat{d}(x,y)=K((x-x_{i}),\sigma_{x}) K((y-y_{i}),\sigma_{y})
\label{kde}
\end{equation}
where the location $(x,y)$ is expressed in bin unit, and where we use the kernel function
\begin{equation}
K(u,\sigma)=\frac{1}{\sigma\sqrt{2\pi}}\exp(\frac{u^2}{2\sigma^2})
\end{equation}
Because the distances are expressed in bin units we scale the sigma's  ($\sigma_{x}$ and $\sigma_{y}$) by dividing the error values $\epsilon$ by the bin size $h$ 
\begin{equation}
\sigma=\frac{\epsilon}{h}
\end{equation}
This implies that the bin size becomes the unit for our density estimates, and we generate probabilities for number of people per bin, which has the area $h\times h$.

>>>>

To visualise the spatiotemporal behaviour of crowd density [in real-time] we dynamically generate heat maps along a moving time interval. 
The data provides us with a series of estimated positions of mobile devices, together with corresponding values of measurement uncertainty (see Section on Positioning).
[The uncertainty values are determined by the co-variance matrix resulting from the chi-square fitting method.]

From the data we select positions that are estimated from measurements that fall within a time window.
From the estimated positions we construct crowd density estimations for the number of mobile devices per unit area.
To visualize the uncertainty that corresponds to each estimated position, we transform it into a bi-variate Gaussian, using the uncertainty values ($\sigma_{x}$ and $\sigma_{y}$).
For each MAC address detected we construct a two-dimensional density estimate that is the sum of these smoothed positions, which we normalize by the number of measurements $N$ for that MAC address.
Thus, the final density estimate for a mobile device is a sum of `bumps' centered at the estimated positions.
Because the number of measurements is different for each detected mobile device, we normalize each histogram separately.
To build our final probability density estimate we sum all the density distributions of all the MAC addresses.
This final density estimate integrates to the number of MAC addresses detected in the time window.

The implementation of our method is similar to kernel density estimation \cite{scott}\cite{silverman}.
In our case the amount of smoothing is variable and determined by the measurement uncertainty values $\sigma_{x}$ and $\sigma_{y}$.

The density estimate for a mobile device $m$ at a location $(x,y)$ is defined by

\begin{equation}
\hat{f}_{m}(x,y)=\frac{1}{N}\sum_{i=1}^{N}K((x-x_{i}),\sigma_{x}) K((y-y_{i}),\sigma_{y})
\label{kde}
\end{equation}
where $N$ is the number of measurements (positions) for MAC address $m$, and the kernel function $K$ is give by
\begin{equation}
K(u,\sigma)=\frac{1}{\sigma\sqrt{2\pi}}\exp(\frac{u^2}{2\sigma^2})
\end{equation}
The final density estimate is given by
\begin{equation}
\hat{f}_{T}(x,y)=\sum_{m}\hat{f}_{m}(x,y)
\label{total}
\end{equation}
To visualize our results as a density histogram in which the estimated values are probabilities for the number of people per bin, we evaluate equation \ref{total} at regular points in a grid, and multiply by the bin size.
